Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2021-12-23 write-up/fda-recurrent.tex
Update fda-recurrent.tex

A key concern is whether all moments of suicidal distress are recorded.  To ensure this, clinical staff interviewed participants in the evening who were then asked to review their button press activity.  Any events that were identified as incorrect button press activity were removed.  At the end of the study period, the average number of button presses per day was $1.91$ with a standard deviation of $2.23$.  Figure~\ref{fig:buttons} presents the distribution of button presses as a function of hour in the day.  This demonstrates low button-press counts prior to 7AM and a sharp drop off by 11PM. This demonstrates an additional concern: events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake. Due to (B), we define each study day to begin at 9AM and end at 8PM. Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.

\vspace{0.25cm}
\begin{minipage}{\textwidth}
  \begin{minipage}[b]{0.49\textwidth}
    \centering
%    \includegraphics[width=0.9\textwidth]{figs/buttonpresses.png}
    \captionof{figure}{Button-presses per hour within the user-day}
    \label{fig:buttons}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\textwidth}
    \centering
    \centering
%    \includegraphics[width=1.0\textwidth]{figs/recurrent_events_plot.png}
    \captionof{figure}{User button-presses versus time in study}
    \label{fig:buttons2}
    \end{minipage}
\end{minipage}
\vspace{0.25cm}

Figure~\ref{fig:buttons2} presents the raw button presses as a function of time in study for each user, showing the potential heterogeneity in button press rates between users and study days. To assess whether there is between user or between study day variation, Table~\ref{tab:anova} presents a two-way ANOVA decomposition of the button press counts as a function of participant and day in study.  The ANOVA decomposition demonstrates high variation with day in study compared to across users in terms of button presses.

\begin{table}[!th]
\begin{tabular}{ccccccc}\hline
      & DF & Sum Sq & Mean Sq & F value & Pr($>$F) \\ \hline
        Participant & 88 & 13 & 13.31 & 2.71 & 0.10 \\
        Day in Study & 75 & 54.71 & 54.71 & $8.8 \times 10^{-4}$ \\
        Residuals & 989 & 4.91 & - \\ \hline
      \end{tabular}  \captionof{table}{ANOVA decomposition of daily button press counts}
      \label{tab:anova}
\end{table}

@@

\subsection{Effect of electrodermal activity and activity index on moments of distress}

Over a period of 5~years and one month following recruitment,
patient~$u$ had eight appointments with prothrombin values as follows:

\section{Discussion}
We abide by the comment of Robert Gentleman on big data: (1) make the data as small as possible as fast as possible. Statistics works!  With subsampling and proper re-weighting the analysis is converted to well-known territory. (2) Make the software and tools as fast and as maintanable as possible.  This is exactly teh benefit of subsampling.  we convert to penalized logistic regression on a smaller but still relatively efficient dataset. This means we can use all the existing tools!

>>>>

A key concern is whether all moments of suicidal distress are recorded.  To ensure this, clinical staff interviewed participants in the evening who were then asked to review their button press activity.  Any events that were identified as incorrect button press activity were removed.  At the end of the study period, the average number of button presses per day was $2.42$ with a standard deviation of $2.62$.  Investigation of the button press data shows low button-press counts prior to 7AM and a sharp drop off by 11PM. This demonstrates an additional concern: events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake. To deal with (A) and (B), here we define each study day to begin at 9AM and end at 8PM.
% Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius.

\vspace{0.25cm}
\begin{figure}[!th]
    \centering
    \includegraphics[width=0.8\textwidth]{../figures/recurrentevents_plot.png}
    \captionof{figure}{User button-presses (red) versus time since study entry (in hours).  The black mark indicates the final sensor measurement time.}
    \label{fig:buttons}
\end{figure}
\vspace{0.25cm}

Figure~\ref{fig:buttons} visualizes button presses versus time since study entry for each user. Here, day 30 is assumed to censor the observation process. A black mark signals dropout before day 30.   Figure~\ref{fig:buttons} shows the potential heterogeneity in button press rates between users and study days. To assess whether there is between user or between study day variation, Table~\ref{tab:anova} presents a two-way ANOVA decomposition of the button press counts as a function of participant and day in study.  The ANOVA decomposition demonstrates high variation with day in study compared to across users in terms of button presses.

\begin{table}[!th]
\begin{tabular}{ccccccc}\hline
      & DF & Sum Sq & Mean Sq & F value & Pr($>$F) \\ \hline
        Participant & 88 & 2675.4 & 30.4 & 8.9 & $< 2 \times 10^{-16}$ \\
        Day in Study & 29 & 443.3 & 15.3 & 4.5 & $3.9 \times 10^{-13}$ \\
        Residuals & 672 & 2304.2 & 3.4 \\ \hline
      \end{tabular}  \captionof{table}{ANOVA decomposition of daily button press counts}
      \label{tab:anova}
\end{table}

@@

% \subsection{Effect of electrodermal activity and activity index on moments of distress}

% Over a period of 5~years and one month following recruitment,
% patient~$u$ had eight appointments with prothrombin values as follows:

\section{Discussion}
We abide by the comment of Robert Gentleman on big data: ``make data as small as possible as fast as possible.'' Statistics works!  With subsampling and proper re-weighting the analysis is converted to well-known territory. (2) Make the software and tools as fast and as maintanable as possible.  This is exactly teh benefit of subsampling.  we convert to penalized logistic regression on a smaller but still relatively efficient dataset. This means we can use all the existing tools!
